---
title: 'Connections: Interfaces to the Outside World'
author: "Gareth R. White"
date: "13/11/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Connections: Interfaces to the Outside World
https://www.coursera.org/learn/r-programming/lecture/pSHGL/connections-interfaces-to-the-outside-world

## Interfaces to the Outside World
Data are read using _connection_ interfaces

- file
- gzfile
- bzfile
- url

## File Connections

```{r}
str(file)
```

- description: name of the file
- open: code indicating "r" read only, "w" writing (and initializing new file), "a" appending, "rb" / "wb", "ab" for binary mode (Windows)

## Connections

```{r}
con <- file("foo.txt", "r")
data <- read.csv(con)
close(con)
# Is all the same as
data <- read.csv("foo.txt")
```


## Reading Lines of a Text File
```{r}
con <- gzfile("words.gz")
x <- readLines(con, 10)
x
```

writeLines takes a character vector, and writes each element one line at at time to a text file.

## Reading Lines of a Text File
Could use for reading webpages.

```{r}
con <- url("http://jhsph.edu", "r")
x <- readLines(con)
head(x)
```

