---
title: "Data Types - Factors"
author: "Gareth R. White"
date: "13 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Types - Factors

## Factors
Used to represent categorical data.
Like an integer vector where each integer has a label.

- Ordered (e.g., something ranked, assistant, associate, full professor.)
- Unordered (e.g., male, female)

Factors are treated specially by modelling functions like lm() and glm()
Using factors with labels is better than using integers.
Factors (labels) are "self-describing".
e.g., "Male" and "Female" are clearer than 1 and 2.

## Factors

```{r}
x <- factor(c("yes", yes", "no", "yes", "no")) # Levels are "yes" and "no"
x
table(x) # Frequency count of each level (e.g., 2 no, and 3 yes)
unclass(x) # Strips class, brings it down to an integer vectos with values 2 2 1 2 1 (where yes is coded as 2 and no is coded as 1) and attribute called levels being "no" and "yes"
```

## Factors
Order of levels can be set by the levels argument to the factor() function.
Baseline level is the first factor.
First factor defaults to first in alphabetic order.

```{r}
x <- factor(c("yes", "yes", "no", "yes", "no"),
      levels = c("yes", "no"))
x
```
