---
title: "Data Types - Matrices"
author: "Gareth R. White"
date: "13 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Types - Matrices
https://www.coursera.org/learn/r-programming/lecture/MzJQR/data-types-matrices


## Matrices
Matrices are a special type of vector, not a separate object, which has an Attribute 'dimension'
dimension is an integer vector of length 2
1st is the number of rows
2nd is number of columns

```{r}
m <- matrix(nrow=2, ncol=3)
m
attributes(m)
```

## Matrices (continued)
constructed columnwise.
i.e., fill down from top left cell, then next column to the right.

```{r}
m <- matrix(1:6, nrow=2, ncol=3)
m
```

## Matrics (cont)

```{r}
m <- 1:10
m
dim(m) <- c(2,5) # 2 rows, 5 columns
m
```

## cbind-ing and rbing-ing
Other ways to create matrices:
-column binding, cbind()
-row binding, rbing()

e.g., column-wise bind them, i.e., the familiar fill down from the top left cell with the first parameter
```{r}
x <- 1:3
y <- 10:12
cbind(x, y)
```

Or row bind, i.e., fill from left-right starting at the top left cell, using the first parameter.
```{r}
rbind(x, y)
```
