---
title: "Data_Types_-_Vectors_and_Lists"
author: "Gareth R. White"
date: "13 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Types - Vectors and Lists 
https://www.coursera.org/learn/r-programming/lecture/wkAHm/data-types-vectors-and-lists

c() function means 'concatenate'
Can create vectors of objects.

```{r}
x <- c(0.5, 0.6) ## numeric
x <- c(TRUE, FALSE) ## logical
x <- c(T,F) ## logical
x <- c("a", "b", "c") ## character
x <- 9:29 ## integer
x <- c(1+0i, 2+4i) ## complex, where the i part represents the imaginary part of the number
```

Or use the vector() function.
Numeric vectors default to being initialised to 0.

```{r}
x <- vector("numeric", length=10)
x
```

## Mixing Objects
R will coerce to the least common denominator.

y <- c(1.7, "a") ## character. 1.7 is coerced to "1.7"
y <- c(TRUE, 2) ## numeric. True is coerced to 1.
y <- c("a", TRUE) ## character. TRUE is coerced to "TRUE"

## Explicit Coercion

Use as.* functions

```{r}
x <- 0:6
class (x) ## integer
as.numeric(x)
as.logical(x)
as.character(x)
as.complex(x)
```

## Explicit Coercion

Nonsense coercions result as NA.

```{r}
x <- c("a", "b", "c")
as.numeric(x)
as.logical(x)
as.complex(x) #0+0i 1+01 2+0i 3+0i 4+0i 5+0i 6+0i ## [GARETH] It actually just produces [1] NA NA NA. Perhaps the slide author made a mistake?
```

## Lists
Can contain elements of different class.
```{r}
x <- list(1, "a", TRUE, 1 + 4i)
x
```

Elements of a list have double brackets around them.
Other vectors have single brackets.


[[1]]
[1] 1
First element is a vector with value 1

[[2]]
[1] "a"
Second element is a vector with value "a"


[[3]]
[1] TRUE
Third element is a vector TRUE

[[4]]
[1] 1+4i
Fourth element is a complex number.
