---
title: "GettingHelp"
author: "Gareth R. White"
date: "13 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Finding Answers
Finding Answers.
Search archives.
Search web.
Read manual.
Read FAQ.
Inspection / experimentation.
Asking a skilled friend.
Review source code.

# Asking Questions
Before asking questions, let others know you've done the earlier steps.

```{r}
library(datasets)
data(airquality)
cor(airquality)
```

https://www.google.com.hk/webhp?sourceid=chrome-instant&ion=1&espv=2&ie=UTF-8#q=missing%20observations%20in%20cov%2Fcor

## Asking Questions
What steps will reproduce the problem?
What is the expected output?
What do you see instead?
What versions are you using (R, packages, OS)

## Subject Headers
"R 3.0.2 lm() function on Mac OS X 10.9.1 -- seg fault on large data frame""

## Do
Describe the goal, not the step.
Be explicit about the question.
Provide the minimum information necessary.
Be courteous.
Follow up with the solution.

## Don't
Claim you've found a bug.
Grovel as a substitute for doing your homework.
Post homework questions in public.
Email multiple mailing lists at once.

## Places t oturn
Class discussion board
r-help@r-project.org

