---
title: "Textual Data Formats"
author: "Gareth R. White"
date: "13/11/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Textual Data Formats
https://www.coursera.org/learn/r-programming/lecture/qFvzp/textual-data-formats

## Textual Formats
dump()
dput()
Text formats.
Editable and potentially recoverable from corruption.
Contain metadata (type of data in each object, class of each column).
Works well with version control systems like git.
Corruption somewhere in the file can be found and fixed easier than binary objects.
Not very space efficient.

## dputing R objets
Will take most R objects.
Constructs R code to recreate the object.
Can only be used on a single R object

```{r}
y <- data.frame(a=1, b=2)
dput(y)
dput(y, file="y.R")
new.y <- dget("y.R")
new.y
```

## Dumping R Objects
Can take multiple objects
Read back using source()
Pass a character vector with the names of the objects.

```{r}
x <- "foo"
y <- data.frame(a=1, b="a")
dump(c("x", "y"), file="data.R")
rm(x,y)
source("data.R")
y
x
```