---
title: "Dates and Times"
author: "Gareth R. White"
date: "20 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dates and Times
https://www.coursera.org/learn/r-programming/lecture/yl7BO/dates-and-times

Date: Date class.
Stored internally as days since 1970.01.01

Time: POSIXct and POSIXlt classes.
Stored internally as seconds since 1970.01.01

## Dates in R

```{r}
x <- as.Date("1970-01-01")
x ## Automatically converted to text
unclass(x) ## 0
unclass(as.Date("1970-01-02")) ## 1
```

## Time in R
POSIXct
Very large integer.
Useful for storing time in data frames.

POSIXlt
List, plus other useful information like day of the week, year, month, etc.

Generic functions for objects of class Date, POSIXct, or POSIXlt:
weekdays() - gives day of the week
months() - gives month name
quarters() - gives quarter number

```{r}
x <- Sys.time()
x
p <- as.POSIXlt(x)
names(unclass(p))
p$sec
```


## Times in R

```{r}
x <- Sys.tim()
x
unclass(x)
x$sec ## Error
p <- as.POSIXlt(x)
p$sec
```

## Times in R
strptime()
Converts text string date/time

```{r}
datestring <- c("January 10, 2012 10:40", "December 9, 2011 9:10")
x <- strptime(datestring, "%B %d, %Y %H:%M")
x
class(x) ## POSIXlt
```

## Operations on Dates and Times
```{r}
x <- as.Date("2012-01-01") ## class Date
y <- strptime("9 Jan 2011 11:34:21", "%d %b %Y %H:%M:%S") ## class POSIXlt
x - y ## Error, can't mix different classes
x <- as.POSIXlt(x)
x - y
```


## Operations on Dates and Times
Keep track of leap years, seconds, daylight savings, timezones.

```{r}
x <- as.Date("2012-03-01")
y <- as.Date("2012-02-28")
x - y # 2 days, because 2012 was a leap year.

x <- as.POSIXct("2012-10-25 01:00:00")
y <- as.POSIXct("2012-10-25 06:00:00", tz = "GMT")
y - x
```

## Summary

- Dates and times have special classes.
- Allow numerical and statistical calculations.
- Dates use _Date_ class.
- Times use _POSIXct_ and _POSIXlt_ classes.
- Character strings can be coerced to Date/time classes using strptime(), as.Date(), as.POSIXct(), or as.POSIXlt()

Plotting functions recognise date/time objects, and format appropriately.
