---
title: "Functions (part 2)"
author: "Gareth R. White"
date: "16 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions (part 2)
https://www.coursera.org/learn/r-programming/lecture/2xbdl/functions-part-2

## Defining a Function

```{r}
f <- function(a, b=1, c=2, d=NULL)
{
}
```

NULL usually means there's nothing there

## Lazy Evaluation

```{r}
f <- function(a, b)
{
  a^2
}
f(2)
```

b is never used.
Arguments to functions are only evaluated as needed.
Calling f(2) without a second parameter is ok.


## Lazy Evaluation

```{r}
f <- function(a, b)
{
  print(a)
  print(b)
}
f(45) ## Error: argument 'b' is missing, with no default
```

## The '...' Argument

Variable number of arguments.
often used when extending another function

```{r}
myplot <- function(x, y, type="1", ...)
{
  plot(x, y, type=type, ...)
}
```

"Generic functions" can pass extra arguments to methods.
Dispatch methods for different types of data.

## The '...' Argument
Also necessary when the number of arguments cnanot be known in advance.
e.g., ... is the first argument in some cases,
args(paste)
args(cat)

## Arguments Coming After the '...' Argument
Must be named.
Cannot be partially matched.

paste("a", "b", sep=":") ## Ok
paste("a", "b", se=":") ## Partial matching of "se" for "sep" does not work.
