---
title: "Functions (part 1)"
author: "Gareth R. White"
date: "15 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions (part 1)

## Functions
Created using the 'function' directive.
Functions are stored as R objects, with class 'function'.
First class objects.
Can be passed as arguments to other functions.
Can be nested. Return value is last expression to be evaluated.

## Function arguments
Named arguments can have default values.
Formal arguments are the arguments included in the function definition.
The 'formals' function returns a list of all the formal arguments of a function.
Not every function call in R makes use of all the formal arguments.

## Argument Matching
Can be matched by position or name.

## Argument Matching
You can mix position and name matching.
Named arguments are matched first.
After matching a name, the rest of the arguments are assigned in order, starting from the first non-named argument.

args(lm)
function (formula, data, subset, weights, na.action,
          method = "qr", model = TRUE, x = FALSE,
          y = FALSE, qr = TRUE, singular.ok = TRUE,
          contrasts = NULL, offset, ...)

The following are equivalent:
```{r}
lm(data = mydata, y ~ x, model = FALSE, 1:100) ## formula is the first non-named argument
lm(y ~ x, mydata, 1:100, model = FALSE) ## forumal is the first non-named argument
```

## Argument Matching
Useful where argument lists are very long.

## Argument Matching
Can be partially matched, for interactive work.
1 Check for exact match.
2 Check for partial match.
3 Check for position match.
