---
title: "Scoping Rules - R Scoping Rules"
author: "Gareth R. White"
date: "16 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Scoping Rules - R Scoping Rules
https://www.coursera.org/learn/r-programming/lecture/FkTgn/scoping-rules-r-scoping-rules

## Lexical Scoping
A function is typically defined in the global environment. Free variables are found in the user's workspace.
This is logical and usually the right thing to do.
However, in R you can define a function inside another function.

```{r}
make.power <- function(n)
{
  pow <- function(x)
  {
    x^n
  }
  pow
}
```

```{r}
cube <- make.power(3)
square <- make.power(2)
cube(3)
square(3)
```

## Exploring a Function Closure

```{r}
ls(environment(cube))
get("n", environment(cube))

ls(environment(square))
get("n", environment(square))
```

## Lexical vs Dynamic Scoping
```{r}
y <- 10
f <- function(x)
{
  y <- 2 ## free
  y^2 + g(x) ## free
}

g <- function(x)
{
  x*y ## y is free
}

f(3) ## g uses y == 10
```

## Lexical Vs. Dynamic Scoping
With lexical scoping, the value y in function g is looked up in the environment in which the function was defined (which here is the global environment).
The value of y is 10.

With dynamic scoping, the value of y is looked for in the environment in which the function was called (the _calling environment_).
In R the calling environment is also known as the _parent frame_.
So the value would be 2.

## Lexical Vs. Dyanamic Scoping
When a function is defined in the global environmetn, and subsequently called from the global environment, then they are the same.
This looks like dynamic scoping (but is not).

```{r}
g <- function (x)
{
  a <- 3
  x+a+y
}

g(2) ## Error, 'y' not found
y <- 3
g(2) ## ok
```


## Other Languages
Other languages support lexical scoping:
- Scheme
- Perl
- Python
- Common Lisp

## Consequence of Lexical Scoping
All objects must be stored in memory.
All functions must carry a poniter to their respective defining environments.
In S-PLUS, free variables were always looked up in the global workspace, so everything can be stored on disk. The defining environment of all functions was the same.

