---
title: "Scoping Rules - Opimization Example (OPTIONAL)"
author: "Gareth R. White"
date: "16 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Scoping Rules - Opimization Example (OPTIONAL)
https://www.coursera.org/learn/r-programming/lecture/Jm1s1/scoping-rules-optimization-example-optional

e.g.,

- optim
- nlm
- optimize

Take a function whos argumetn is a vector of parameters (e.g., a log-likelihood)

NOTE: Optimization is explained quite simply here, http://www.purplemath.com/modules/linprog.htm


## Maximizing a Normal Likelihood

Write a "constructor" function to create the "objective"" function

```{r}
make.NegLogLik <- function(data, fixed=c(FALSE, FALSE))
{
  params <- fixed
  function(p)
  {
    params[!fixed] <- p
    mu <- params[1]
    sigma <- params[2]
    a <- -0.5*length(data) * log(2*pi*sigma^2)
    b <- -0.5*sum((data-mu)^2) / (sigma^2)
    -(a+b)
  }
}
```

Optimization functions in R _minimize_ functions, so you need to use the negative log-likelihood.

## Maximizing a Normal Likelihood

```{r}
set.seed(1); normals <- rnorm(100, 1, 2)
nLL <- make.NegLogLik(normals)
nLL
ls(environment(nLL))
```

## Estimating Parameters

```{r}
optim(c(mu = 0, sigma = 1), nLL)$par

# optimize() is a 1D optimzation, only takes one value
# so in function(p), params[!fixed] <- p, is assigning a 1D vector (the current value to test, passed from the optimze() function) to a 1D vector params[1] in this case, which is mu, while params[2] is fixed to the value 2 and the optimize() function does not modify it.
nLL <- make.NegLogLik(normals, c(FALSE, 2)) # sigma = 2
optimize(nLL, c(-1, 3))$minimum

nLL <- make.NegLogLik(normals, c(1, FALSE)) # mu = 1
optimize(nLL, c(1e-6, 10))$minimum
```


## Plotting the Likelihood

```{r}
nLL <- make.NegLogLik(normals, c(1, FALSE))
x <- seq(1.7, 1.9, len=100)
y <- sapply(x, nLL)
plot(x, exp(-(y - min(y))), type = "1")

nLL <- make.NegLogLik(normals, c(FALSE, 2))
x <- seq(0.5, 1.5, len = 100)
y <- sapply(x, nLL)
plot(x, exp(-(y - min(y))), type = "1")
```
