---
title: "Scoping Rules - Symbol Binding"
author: "Gareth R. White"
date: "16 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Scoping Rules - Symbol Binding

## A Diversion on Binding Values to Symbol
Search through a series of _environments_

At command line:
- Global environment for a symbol matching the one requested
- Namespaces of each package in the search list.

Search lists are shown using _search()_
This is an ordered list.
".GlobalEnv" is always first.

## Binding Values to Symbol
Global environment is always first.
Base package is always last.
The order matters.
Users can configure which get loaded at startup.
When using _library()_, the namespace gets put in position 2 (by default).
R has separate namespaces for functions and non-functions. It's possible to have an object named _c_ and a function named _c()_

## Scoping Rules
R uses _lexical scoping_ (aka _static scoping_). A common alternative is dynamic scoping.

## Lexical Scoping

```{r}
f <- function(x, y)
{
  x^2 + y / z
}
```

z is a _free variable_ not deined in the function header.

## Lexical Scoping
"the values of free variables are searched for in the environment in which the function was _defined_"

An _environment_ is a collection of (symbol, value) pairs.
Every environment has a parent environment, and can have multiple children.
The _empty_ environment has no parent.
A function + environment is a _closure_ or _function closure_

## Lexical Scoping
If the value of a symbol is not found in the environment in which a function was defined, then the search is continued in the parent environment.
Search continues up through the parent envrionments until we hit the _top-level_ environment (e.g., the global environment - the workspace - or the namespace of the package.
After the top-level environment, search continues until we hit the _empty_ environment.
