---
title: "apply"
author: "Gareth R. White"
date: "21 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# apply
https://www.coursera.org/learn/r-programming/lecture/IUUhK/loop-functions-apply\

To evaluate a function over the margins of an array.
Often uses anonymous functions.
Most often used over the rows or columns of a matrix.
Can be used with general arrays (e.g., average of an array of matrices).
Generally _not_ faster than writing a loop.
Code looks more compact (e.g., 1 line, instead of multi-line for-loop).

## apply
- x, is an array (array is vector that has dimensions).
- MARGIN, is an integer vector, indicating which margins should be 'retained'.
- FUN, is the function to apply.
- ..., is other arguments passed to FUN.

## apply

```{r}
x <- matrix(rnorm(200), 20, 10) # 20 rows, 10 columns, elements random
apply(x, 2, mean)
# MARGIN = 2 means retain the 2nd dimension (columns), and collapse the 1st (rows).
# Returns vector of length 10:
# The means of all 20 rows, for each of the 10 columns.

apply(x, 1, sum)
# MARGIN = 1 means retain the 1st dimension (rows), and collapse the 2nd (columns).
# Returns vector of length 20:
# The sum of all 10 columns, for each of the 20 rows.
```

## col / row sums and means
Much faster shortcuts:

- rowSums() = apply(x, 1, sum)
- rowmeans() = apply(x, 1, mean)
- colSums() = apply(x, 2, sum)
- colMeans() = apply(x, 2, mean)

## Other ways to apply

Quantiles of the rows of a matrix.
```{r}
x <- matrix(rnorm(200), 20, 10)
apply(x, 1, quantile, probs = c(0.25, 0.75))
# MARGIN = 1, preserve rows.
# probs = 25th and 75th percentile
# For each row, calculate 2 numbers: the 25th and 75th percentile.
# Returns 2x20 matrix, where row 1 is the 25th percentile, row 2 is the 75th percentile.
# 20 columns, corresponding to the 20 preserved rows of x.
```


## apply
Average matrix in an array

```{r}
a <- array(rnorm(2 * 2 * 10), c(2, 2, 10))
## 3D array, with dimensions 2, 2, 10.
## Filled with random variables

apply(a, c(1, 2), mean)
## MARGIN = 1st and 2nd dimensions (collapse 3rd)
## Returns a matrix with dimensions 2x2, where each element is the mean of the 10 elements of the 3rd dimension of a

rowMeans(a, dims = 2)
## Returns the same thing.
```
