---
title: "Loop Functions - mapply"
author: "Gareth R. White"
date: "21 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loop Functions - mapply
https://www.coursera.org/learn/r-programming/lecture/EBnAr/loop-functions-mapply

## mapply
Multivariate version of lapply / sapply.
Applies a function _in parallel_ over a _set_ of arguments.
apply only applies a function to a single argument, which is a single list, not to multiple list arguments.

- FUN is the function to apply. Number of arguments has to be at least as many as the number of lists you pass to mapply.
- ... contains arguments to apply over.
- MoreArgs is a list of other arguments to pass to FUN.
- SIMPLIFY indicates whether the result should be simplified.

## mapply

These 2 lines are equivalent:
```{r}
list(rep(1, 4), rep(2, 3), rep(3, 2), rep(4, 1))
## Returns a list of 4 elements:
	## List of 4 elements: 1111
	## List of 3 elements: 222
	## List of 2 elements: 33
	## List of 1 elements: 4

mapply(rep, 1:4, 4:1)
## Invokes rep 4 times: (1,4), rep(2, 3), rep(3, 2), rep(4, 1)
## Returns the same
```

_rep_ replicates.
