---
title: "Loop Functions - mapply"
author: "Gareth R. White"
date: "21 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loop Functions - mapply
https://www.coursera.org/learn/r-programming/lecture/EBnAr/loop-functions-mapply

## mapply
Multivariate version of lapply / sapply.
Applies a function _in parallel_ over a _set_ of arguments.
apply only applies a function to a single argument, which is a single list, not to multiple list arguments.

- FUN is the function to apply. Number of arguments has to be at least as many as the number of lists you pass to mapply.
- ... contains arguments to apply over.
- MoreArgs is a list of other arguments to pass to FUN.
- SIMPLIFY indicates whether the result should be simplified.

## mapply

These 2 lines are equivalent:
```{r}
list(rep(1, 4), rep(2, 3), rep(3, 2), rep(4, 1))
## Returns a list of 4 elements:
	## List of 4 elements: 1111
	## List of 3 elements: 222
	## List of 2 elements: 33
	## List of 1 elements: 4

mapply(rep, 1:4, 4:1)
## Invokes rep 4 times: (1,4), rep(2, 3), rep(3, 2), rep(4, 1)
## Returns the same
```

_rep_ replicates the first argument, a number of times specified by the second argument.

## Vectorizing a Function

```{r}
noise <- funciton(n, mean, sd)
{
	rnorm(n, mean, sd)
}

noise(5, 1, 2)

noise(1:5, 1:5, 2)
# Doesn't really vectorize as you might hope:
# n is 1:5, which in this case is interpreted by rnorm as length(1:5), therefore n=5
# Each observation takes a different mean, from 1:5
```

## Instant Vectorization
Useful for vectorizing a function that doesn't allow for vectors of arguments.

```{r}
mapply(noise, 1:5, 1:5, 2)
## Invokes noise 5 times.
## n=1, mean=1, sd=1
## n=2, mean=2, sd=1
## n=3, mean=3, sd=1
## n=4, mean=4, sd=1
## n=5, mean=5, sd=1
```

## Instant Vectorization
This is the same as,
```{r}
list(noise(1, 1, 2), noise(2, 2, 2), noise(3, 3, 2), noise(4, 4, 2), noise(5, 5, 2))
```

}