---
title: "Loop Functions - tapply"
author: "Gareth R. White"
date: "22 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loop Functions - tapply
https://www.coursera.org/learn/r-programming/lecture/w98BR/loop-functions-tapply

## tapply
Apply a function over a _subset of a vector_.
e.g., calculate a mean for different groups in the vector.

- _x_ is a vector.
- _INDEX_ is a factor, or a list of factors (coerced to factors if necessary). Same length as x
- _FUN_ is the function to apply.
- _..._ other arguments passed to FUN.
- _simplify_ indicates if the result should be simplified, like sapply.

## tapply
```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1)) # Vector of 30 elements.
f <- gl(3, 10) # Generate a factor variable with 3 levels, with 10 replications (i.e., f is a vector of 30 elements, first 10 are 1, second 10 are 2, 3rd 10 are 3)
f
tapply(x, f, mean) # Computes mean for 3 sets (with indices given by f) of 10 elements (from x)
# Results in an array of 3 numerics, labeled with the factor names.
```

## tapply
Take group means without simplification.

```{r}
tapply(x, f, mean, simplify = FALSE)
# Returns a list
```

## tapply
Find group ranges.

```{r}
tapply(x, f, range)
# Returns a list of 3 elements (named by f), where each is a vector of 2 numerics, indicating the min and max of the group.
```
