---
title: "Loop Functions - lapply"
author: "Gareth R. White"
date: "21 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loop Functions - lapply
https://www.coursera.org/learn/r-programming/lecture/t5iuo/loop-functions-lapply

## Looping on the Command Line

lapply - loop over list, and apply a function to every element.
sapply - simplifies the result
apply - over the margins of an array
tapply - over subsets of a vector
mapply - multivariate variation of lapply

spliy is useful in conjunction with lapply to split objects into sub pieces.


## lapply

1. x is a list (coerced if necessary)
2. FUN is a function
3. other arguments ...

## lapply
Always returns a list, but not necessarily the same type of list.

```{r}
x <- list(a = 1:5, b = rnorm(10)) # element a has 5 items, element b has 10.
lapply(x, mean) # returned list has 2 elements, the mean for a and the mean for b
```

## lapply
```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
lapply(x, mean)
```

## lapply
```{r}
x <- 1:4
lapply(x, runif)
# runif generates random deviates, where it's first parameter n is the number of observations required.
# runif is invoked 4 times, once per element in the sequence 1, 2, 3, 4.
#i.e., this produces 4 deviates, with 1, 2, 3, and 4 observations respectively.
```

## lappy
```{r}
x <- 1:4
lapply(x, runif, min = 0, max = 10)
```

## lapply
The *apply functions often use _anonymous_ functions.

```{r}
x <- list(a = matrix(1:4, 2, 2), b = matrix(1:6, 3, 2))
x
```

An anonymous function to extract the first column from each matrix:

```{r}
lapply(x, function(elt) elt[,1])
```

## sapply

- If the result is a list where every element is length 1, then a vector is returned
- If the result is a list where every element is a vector of the same length (>1), a matrix is returned.
- Otherwise a list is returned.

### sapply

```{r}
x <- list(a = 1:4, b = rnorm(10), c = rnorm(20, 1), d = rnorm(100, 5))
lapply(x, mean) # Returns list of length 4, where each element is a vector of length 1, where each item is numeric
sapply(x, mean) # Returns vector of length 4, where each element is a numeric
mean(x) ## error, can't apply to lists - need to use *apply
```
