---
title: "Loop Functions - split"
author: "Gareth R. White"
date: "22 November 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loop Functions - split
https://www.coursera.org/learn/r-programming/lecture/2VYGZ/loop-functions-split

## Split
Takes a vector of other objects, and splits it into groups indicated by a factor or list of factors.
Like tapply, but without FUN.
After splitting you could use lapply or sapply to each separate group.

- _x_ is a vector (or list) or data frame.
- _f_ is a factor (or coerced to one) or a list of factors.
- _drop_ indicates whether empty factors should be dropped.

## Split
```{r}
x <- c(rnorm(10), runif(10), rnorm(10, 1))
f <- gl(3, 10)
split(x, f)
# Returns a list of 3 vectors, each with 10 numeric elements.
```

## Split
Common to use split, followed by lapply.
```{r}
lapply(split(x, f), mean)
# Equivalent to tapply
```

## Splitting a Data Frame
```{r}
library(datasets)
head(airquality)
```

## Splitting a Data Frame
Calculate separate means, for each month.
Split the data frame into monthly pieces.

```{r}
s <- split(airquality, airquality$Month) # Month is coerced into a factor variable
lapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")])) # Get 3 means for each month.
# Returns a list, with the number of elements indicated by the number of months.
# Each element is a vector of 3 numeric means, where each element has 3 names() corresponding to the month it represents.
```

## Splitting a Data Frame
```{r}
sapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Wind")]))
# Returns 2d matrix, of 3 reading type rows x 5 month columns.
sapply(s, function(x) colMeans(x[, c("Ozone", "Solar.R", "Windd")], na.rm = TRUE)) # Remove NAs before calculating column mean
```

## Splitting on More than One Level
```{r}
x <- rnorm(10)
f1 <- gl(2, 5) # e.g., gender
f2 <- gl(5, 2) # e.g., race
f1
f2
interaction(f1, f2)
# Computes a factor, that represents the interaction of the given factors.
# 2 gender x 5 race = 10 new factors.
```

## Splitting on More than One Level
Interactions can create empty levels.
```{r}
str(split(x, list(f1, f2))) # Automatically calls the interaction function to create the new factor variable representing the 10 interaction factor levels of f1 and f2.
# Returns list of 10 (interaction factor levels), where each element is a numeric vector
```

## split
Empty levels can be dropped.
```{r}
str(split(x, list(f1, f2), drop = TRUE))
# Returns a list of 6 elements, which are vectors of numerics.
```